The human in the loop: volunteer-based metacomputers as a socio-technical system

### 2016-04-24 Alife16/paper.tex
Changed all to socio-technical to be coherent with the title closes #119

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other
resources they will ``lend''. This makes them an essential part of the
algorithm and of the performance of the whole system. As a
sociotechnical system, this participation follows some patterns and in
this paper we examine the result of several volunteer distributed
evolutionary computation experiments and try to find out what
those patterns are and what makes an experiment successful or not,
including the feedback loop that is created between the users and the
algorithm itself.
\end{abstract}

@@

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island carrying
out their own evolutionary algorithms. This is why NodIO is proposed
mainly as a {\em complement} to existing resources such as desktop
systems or laptops. As long as it provides a non-null computational
capability that can help existing resources find the solution faster
it will have found its purpose. The main use case is someone setting
up NodIO in the cloud, writing a fitness evaluation function and
running a client from his or her own computer, but requesting help in
social networks for additional resources. That is why, in this system,
we have considered the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. The
computing system becomes a {\em sociotechnical system}
\citep{vespignani2009predicting}. In this paper, we are going to focus
on measuring the response of users to experiments, that is, the time
they spend running it, but at the same time we will also focus on the
technical aspects of the server and how these might change the
behavior of users, improving the capability of the system.

@@

As far as we know, this paper presents one of the few experiments
that measure the performance of a sociotechnical
metacomputer, that is, a spontaneously created parallel computer that
uses social networks for operations such as gathering new
users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the 
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. As in the previous example, a social network
was used to get new network nodes; in the previous case a web page was
used, while Facebook's wall was used here.

@@

These experiments show that, as it was proved for other volunteer
computer frameworks and also in the case of games, user engagement
follows a Weibull distribution. This makes engagement the key for
leveraging the performance of the sociotechnical metacomputer and a
way to improve results in the future.

@@

Our intention in this paper was to assess the capabilities of a
sociotechnical system formed by a client-server web-based framework
running a distributed evolutionary algorithms and the volunteers that
participate in the experiment. These volunteers are {\em in the
  cloud}, that is, available as {\em CPU as a service} for the persons
running the experiment. In this paper we have tried to put some figures
on the real size of that {\em cloud} and how it can be used standalone
if there is no alternative, or, if other computing resources are
available, in conjunction with other local or cloud-based methods to
add computing power in a seamless way through the pool that NodIO creates.

After running the experiment on the 40-trap problem whose running time
could be as low as a few seconds, we switched to another batch of
experiments where we used the 50-trap problem and also to a pool of
limited, and dwindling through the three experiments, size. Since our
initial results indicated that what happened on the screen, a flat
graph with no improvements or the experiment finished, influenced the
amount of time that the users devoted to the experiment, a longer one
could yield different results and, at the same time, result in a big
pool that might either crash the server or return useless individuals
to the volunteers. These new experiments have proved that using the
limited pool is beneficial to finding the solution, since less
evaluations are needed, but also that since the problem is more
difficult, the users stay for longer in the web page, making less
ephemeral the sociotechnical computing system created by the
simulation.

>>>>

\begin{abstract}
Volunteer computing is a form of distributed computing where users
decide on their participation and the amount of time and other
resources they will ``lend''. This makes them an essential part of the
algorithm and of the performance of the whole system. As a
socio-technical system, this participation follows some patterns and in
this paper we examine the result of several volunteer distributed
evolutionary computation experiments and try to find out what
those patterns are and what makes an experiment successful or not,
including the feedback loop that is created between the users and the
algorithm itself.
\end{abstract}

@@

This loose connection provides a low-overhead way to connect desktop
experiments with volunteer-based ones, with all of them contributing
to the pool, but every one of them working as separate island carrying
out their own evolutionary algorithms. This is why NodIO is proposed
mainly as a {\em complement} to existing resources such as desktop
systems or laptops. As long as it provides a non-null computational
capability that can help existing resources find the solution faster
it will have found its purpose. The main use case is someone setting
up NodIO in the cloud, writing a fitness evaluation function and
running a client from his or her own computer, but requesting help in
social networks for additional resources. That is why, in this system,
we have considered the whole {\em social} aspect in the design, with
issues related to security, trust and privacy among others. The
computing system becomes a {\em socio-technical system}
\citep{vespignani2009predicting}. In this paper, we are going to focus
on measuring the response of users to experiments, that is, the time
they spend running it, but at the same time we will also focus on the
technical aspects of the server and how these might change the
behavior of users, improving the capability of the system.

@@

As far as we know, this paper presents one of the few experiments
that measure the performance of a socio-technical
metacomputer, that is, a spontaneously created parallel computer that
uses social networks for operations such as gathering new
users. Apol{\'o}nia et al. \citep{apolonia2012enhancing} used the 
Facebook protocol to distribute tasks among the {\em walls} of
friends, explicitly using the social network for computing. However,
it stopped short of relating performance to the macro measures of the
users' social networks. As in the previous example, a social network
was used to get new network nodes; in the previous case a web page was
used, while Facebook's wall was used here.

@@

These experiments show that, as it was proved for other volunteer
computer frameworks and also in the case of games, user engagement
follows a Weibull distribution. This makes engagement the key for
leveraging the performance of the socio-technical metacomputer and a
way to improve results in the future.

@@

Our intention in this paper was to assess the capabilities of a
socio-technical system formed by a client-server web-based framework
running a distributed evolutionary algorithms and the volunteers that
participate in the experiment. These volunteers are {\em in the
  cloud}, that is, available as {\em CPU as a service} for the persons
running the experiment. In this paper we have tried to put some figures
on the real size of that {\em cloud} and how it can be used standalone
if there is no alternative, or, if other computing resources are
available, in conjunction with other local or cloud-based methods to
add computing power in a seamless way through the pool that NodIO creates.

After running the experiment on the 40-trap problem whose running time
could be as low as a few seconds, we switched to another batch of
experiments where we used the 50-trap problem and also to a pool of
limited, and dwindling through the three experiments, size. Since our
initial results indicated that what happened on the screen, a flat
graph with no improvements or the experiment finished, influenced the
amount of time that the users devoted to the experiment, a longer one
could yield different results and, at the same time, result in a big
pool that might either crash the server or return useless individuals
to the volunteers. These new experiments have proved that using the
limited pool is beneficial to finding the solution, since less
evaluations are needed, but also that since the problem is more
difficult, the users stay for longer in the web page, making less
ephemeral the socio-technical computing system created by the
simulation.
